---
title: "Assignment 7"
author: 'Terrance Liang 40158057'
output: 'pdf_document'
---

## Chapter 5.2 question 6

a)
```{r}
# run the random number generation
u1 <- runif(10, 0, 1)
u2 <- runif(10, 0, 1)
# computing the expected value in the indicated way
mean(u1 + u2)
mean(u1) + mean(u2)
```
We see that the the results of the calculation of the two expected values are the same. Additonally, we know 
that both random variables are uniformly distributed, thus they're expected value is (a+b)/2, thus each random variable 
has an expected value of 1/2 making the sum of the random variables 1.

b)

```{r}
u1 <- runif(1000, 0, 1)
u2 <- runif(1000, 0, 1)
var(u1 + u2)
var(u1) + var(u2)
```
The values are not equal. By mathematical statistics by wackerly, we know the variance of the sum
of random variables is var(u1) + var(u2) + 2cov(u1,u2). Thus the equality proposed is only true if both random 
variables are independent making the covariance 0. In our problem it is said that our random variables are approx. 
independent, thus the covariance should not be 0 and the equations should not be exactly equal. However, in theory, if 
the two rv's are indepedent, then the equality should have been equal since the covariance would have been 0.

c)

```{r}
u1 <- runif(1000, 0, 1)
u2 <- runif(1000, 0, 1)
# we count the number of elements in u1+u2 st. the inequality is true
# this will create a true false vector for which we can pick the # of elements 
# which respectecs the condition. we could have picked u1,u2 or from u1+u2, it does not matter
# since we only want the length to be the same.
length(u1[(u1 + u2 <= 1.5)]) / length(u1)
```

d)

```{r}
u1 <- runif(1000, 0, 1)
u2 <- runif(1000, 0, 1)
# we count the number of elements in u1+u2 st. the inequality is true
# this will create a true false vector for which we can pick the # of elements 
# which respectecs the condition. we could have picked u1,u2 or from u1+u2, it does not matter
# since we only want the length to be the same.
length(u1[(sqrt(u1) + sqrt(u2) <= 1.5)]) / length(u1)
```

## Chapter 5.2 question 8

```{r}
discreterunif <- round(runif(1000, 1, 10))
table(discreterunif)
```

The distribution should be uniform, however it appears that the tails have fewer frequency that the other 
values. A better procedure to keep the uniform distribution would be to use a ceiling function.

```{r}
discreterunif <- ceiling((runif(1000, 1, 10)))
table(discreterunif)
```

## Chapter 5.2 question 10

a)

```{r}
U2 <- runif(1000)
U1 <- runif(1000)
X <- U1 + U2
Y <- U1 - U2
plot(Y ~ X)
cor(X, Y)
```

No they are not linearly dependent, the corelation value is very close to 0 which indicates that there is 
no linear depedence between the two random variables.

b)

They are not stochastically independent because there is a pattern in the plot. This means that there is some 
correlation between x and y.

c) 

```{r}
U2 <- runif(1000)
U1 <- runif(1000)
plot(U1 ~ U2)
cor(U1, U2)
```

we see that the correlatio value is also very close to 0, thus they are not linearly dependent.

d)

They are stochastically independent because there is no discernable pattern in the plot, which shows
that the random variables are no dependent with each other.

## Chapter 5.3.1 question 2

a)
```{r}
# create a function which simulates a bernoulli trial of 20 questions
true_false <- function(n) {
    rbinom(n, size = 1, prob = 0.5)
}

# compute the average for each student on 20 questions 100 times(20 times for every student)
vector <- c()
for (i in (1:100)) {
    temp <- true_false(20)
    vector <- append(vector, length(temp[temp == 1]) / 20)
}

mean(vector)
sqrt(var(vector))
```

b)

```{r}
true_false <- function(n) {
    rbinom(n, size = 1, prob = 0.5)
}
vector <- c()
for (i in (1:100)) {
    temp <- true_false(20)
    vector <- append(vector, length(temp[temp == 1]) / 20)
}

# length of values above 0.3 divided by total number of values
length(vector[vector >= 0.3]) / length(vector)
```

## Chapter 5.3.2 question 3

```{r}
binomial <- rbinom(500, size = 1, prob = 0.99)
mean(binomial)
var(binomial)

0.99
0.99 * 0.01
```
The expectation and variance of a bernoulli distribution by mathematical statistics is respectively p and p*q

## Chapter 5.3.2 question 7

a)

```{r}
par(mfrow = c(10, 10), mar = c(1, 1, 1, 1))

for (m in 1:100) {
    z <- (rbinom(20000, size = m, prob = 0.4) - m * 0.4) /
        sqrt(m * 0.4 * 0.6)
    qqnorm(z, ylim = c(-4, 4), main = paste("QQ-plot, m = ", m))
    qqline(z)
}
```

we see that as m increases the line is slowly getting more and more linear, thus as m
increases the distribution approaches a normal distribution.

b)

Modified code with 100 iterations for each p value
```{r}
# par(mfrow = c(10, 10), mar = c(1, 1, 1, 1))

# for (m in 1:100) {
#     z <- (rbinom(20000, size = m, prob = 0.3) - m * 0.4) /
#         sqrt(m * 0.4 * 0.6)
#     qqnorm(z, ylim = c(-4, 4), main = paste("QQ-plot, m = ", m))
#     qqline(z)
# }

# for (m in 1:100) {
#     z <- (rbinom(20000, size = m, prob = 0.2) - m * 0.4) /
#         sqrt(m * 0.4 * 0.6)
#     qqnorm(z, ylim = c(-4, 4), main = paste("QQ-plot, m = ", m))
#     qqline(z)
# }

# for (m in 1:100) {
#     z <- (rbinom(20000, size = m, prob = 0.1) - m * 0.4) /
#         sqrt(m * 0.4 * 0.6)
#     qqnorm(z, ylim = c(-4, 4), main = paste("QQ-plot, m = ", m))
#     qqline(z)
# }

# for (m in 1:100) {
#     z <- (rbinom(20000, size = m, prob = 0.05) - m * 0.4) /
#         sqrt(m * 0.4 * 0.6)
#     qqnorm(z, ylim = c(-4, 4), main = paste("QQ-plot, m = ", m))
#     qqline(z)
# }
```

```{r}
par(mfrow = c(5, 5), mar = c(1, 1, 1, 1))

for (m in 1:25) {
    z <- (rbinom(20000, size = m, prob = 0.3) - m * 0.4) /
        sqrt(m * 0.4 * 0.6)
    qqnorm(z, ylim = c(-4, 4), main = paste("QQ-plot, m = ", m))
    qqline(z)
}
```
A stright line is apparent are m = 21
```{r}
par(mfrow = c(5, 5), mar = c(1, 1, 1, 1))

for (m in 1:25) {
    z <- (rbinom(20000, size = m, prob = 0.2) - m * 0.4) /
        sqrt(m * 0.4 * 0.6)
    qqnorm(z, ylim = c(-4, 4), main = paste("QQ-plot, m = ", m))
    qqline(z)
}
```
After trying the first 25 iterations of m with no complete straight line, I am 
trying the next 25 to 50 for which m=26 gives a fairly sraight line
```{r}
par(mfrow = c(5, 5), mar = c(1, 1, 1, 1))

for (m in 26:50) {
    z <- (rbinom(20000, size = m, prob = 0.1) - m * 0.4) /
        sqrt(m * 0.4 * 0.6)
    qqnorm(z, ylim = c(-4, 4), main = paste("QQ-plot, m = ", m))
    qqline(z)
}
```
After trying the first 25 iterations of m with no complete straight line, I am 
trying the next 25 to 50 for which m=43 gives a fairly sraight line
```{r}
par(mfrow = c(5, 5), mar = c(1, 1, 1, 1))

for (m in 76:100) {
    z <- (rbinom(20000, size = m, prob = 0.05) - m * 0.4) /
        sqrt(m * 0.4 * 0.6)
    qqnorm(z, ylim = c(-10, 0), main = paste("QQ-plot, m = ", m))
    qqline(z)
}
```
After trying the first 75 iterations of m with no complete straight line, I am 
trying the next 75 to 100 for which m=100 gives the best straight line, a better
straight line can be obtained after 100 iterations

b)
No it is not, for p = 0.05 more iterations are needed to obtain a better straight line
with the extremities not deviating from the straight line. But in the other 3 cases, yes it is 
satisfactory. We see that as p decreases more iterations are needed.

## Chapter 5.3.3 question 6

```{r}
par(mfrow = c(8, 8), mar = c(1, 1, 1, 1))
for (m in seq(1, 120, 2)) {
    z <- (rpois(20000, lambda = m) - m) / sqrt(m)
    qqnorm(z, ylim = c(-4, 4), main = "QQ-plot")
    qqline(z)
    mtext(bquote(lambda == .(m)), 3)
}
```
as lambda increases we see that the plot approaches a normal distribution from
how the dots are alligning with the straight line.

b)

A straight line can be observed after lambda = 21 in the QQ plot without any
of the tails deviating from the straight line.

## Chapter 5.3.3 question 7

a)

```{r}
# generate 10000 poisson pseudorandom numbers with parameter 2.5 x 2
N <- rpois(10000, lambda = 2.5 * 2)
points <- list()

# for each  poisson value generated, we generate N of it;s uniform psudorandom numbers
# on interval 0 to 2
for (i in 1:10000) {
    points[[i]] <- runif(N[i], 0, 2)
}

# counting the points from 1 to 2
vector1 <- c()
for (i in 1:10000) {
    vector1 <- append(vector1, length(points[[i]][points[[i]] > 1]))
}
vector1[(1:15)]

# counting the points from 0 to 1
vector2 <- c()
for (i in 1:10000) {
    vector2 <- append(vector2, length(points[[i]][points[[i]] < 1]))
}
vector2[(1:15)]

length(points[[1]])
```
we only output the first 15 values since there are too many values, the two vector's first value should sum to the length of points[[1]]
b)

```{r}
table(vector1)
table(vector2)
```

Yes the distributions are similar, thus every interval follows a poisson distribution with lambda 2.5

c)

```{r}
plot(jitter(vector1), jitter(vector2))
plot(vector1,vector2)
```

Jitter adds a small amount of noise to the numeric vector. This helps in our situation
because there is a lot of overlap in the points, making them indiscernable.

Because there is no pattern, there is no dependence between the two ranges of points.