---
title: "Assignment 5"
author: 'Terrance Liang 40158057'
output: 'pdf_document'

---

## Chapter 4.1.1 question 4

```{r}
# nested for loop to to try for the three cases
for (x in c(0.006, 0.005, 0.004)) {
    i <- x
    for (j in 1:20) {
        i <- (1 - (1 + i)^(-20)) / 19
    }
    print(i)
}
# computed the mean of the three cases and rounded to 5 digits
round(mean(c(0.005264348, 0.00495812, 0.004558684)), digits = 5)

for (x in c(0.006, 0.005, 0.004)) {
    i <- x
    for (j in 1:1000) {
        i <- (1 - (1 + i)^(-20)) / 19
    }
    print(i)
}
```
we see that the mean using 20 iterations is 0.00493, we confirm that it approches that number by performing 
1000 iterations in those same starting numbers.

## Chapter 4.1.1 question 5

```{r}
for (i in c(2, 3, 4, 5)) {
    x <- 0.5
    pointer <- 0
    # while condition that keeps only the right amount of significant digits, 
    # and if their subtraction is not 0 it keeps iterating
    while (abs(signif(x, i) - signif(cos(x), i)) != 0) {
        # assigns the equation to the x to compare with the next value
        x <- cos(x)
        # counts the number of iterations
        pointer <- pointer + 1
    }
    print(pointer)
}
print("")
for (i in c(2, 3, 4, 5)) {
    x <- 0.7
    pointer <- 0
    while (abs(signif(x, i) - signif(cos(x), i)) != 0) {
        x <- cos(x)
        pointer <- pointer + 1
    }
    print(pointer)
}
print("")
for (i in c(2, 3, 4, 5)) {
    x <- 0
    pointer <- 0
    while (abs(signif(x, i) - signif(cos(x), i)) != 0) {
        x <- cos(x)
        pointer <- pointer + 1
    }
    print(pointer)
}
```
We notice that as the tolerance gets smaller, more iterations are needed, which is expected since to obtain a more precise estimate we need more iterations.
We also notice that as our starting point becomes bigger, there are less iterations needed, thus our solution is closer to 0.7 than it is to the other points.

## Chapter 4.1.1 question 6.a)

```{r}
for (i in c(2, 3, 4, 5)) {
    x <- 0.5
    pointer <- 0
    while (abs(signif(x, i) - signif(1.5 * cos(x), i)) != 0) {
        x <- 1.5 * cos(x)
        pointer <- pointer + 1
        if (pointer > 1000000) {
            print(pointer)
            break
        }
    }
    print(pointer)
}
print("")
for (i in c(2, 3, 4, 5)) {
    x <- 0.7
    pointer <- 0
    while (abs(signif(x, i) - signif(1.5 * cos(x), i)) != 0) {
        x <- 1.5 * cos(x)
        pointer <- pointer + 1
        if (pointer > 1000000) {
            print(pointer)
            break
        }
    }
    print(pointer)
}
print("")
for (i in c(2, 3, 4, 5)) {
    x <- 0
    pointer <- 0
    while (abs(signif(x, i) - signif(1.5 * cos(x), i)) != 0) {
        x <- 1.5 * cos(x)
        pointer <- pointer + 1
        if (pointer > 1000000) {
            print(pointer)
            break
        }
    }
    print(pointer)
}
```

For 1.5cos(x) none of the iterations for starting points 0, 0.5, 0.7 converges after 1 million iterations
We know from numerical analysis that |g'(x)| needs to be smaller than 1 for x e [0,1], if we use x as 0.9149 we obtain 
that |g'(x)| = 1.18875 which means it diverges.
```{r}
for (i in c(2, 3, 4, 5)) {
    x <- 0.5
    pointer <- 0
    while (abs(signif(x, i) - signif((cos(x) / 30) + (44 * x / 45), i)) != 0) {
        x <- (cos(x) / 30) + (44 * x / 45)
        pointer <- pointer + 1
        if (pointer > 1000000) {
            print(pointer)
            break
        }
    }
    print(pointer)
}
print("")
for (i in c(2, 3, 4, 5)) {
    x <- 0.7
    pointer <- 0
    while (abs(signif(x, i) - signif((cos(x) / 30) + (44 * x / 45), i)) != 0) {
        x <- (cos(x) / 30) + (44 * x / 45)
        pointer <- pointer + 1
        if (pointer > 1000000) {
            print(pointer)
            break
        }
    }
    print(pointer)
}
print("")
for (i in c(2, 3, 4, 5)) {
    x <- 0
    pointer <- 0
    while (abs(signif(x, i) - signif((cos(x) / 30) + (44 * x / 45), i)) != 0) {
        x <- (cos(x) / 30) + (44 * x / 45)
        pointer <- pointer + 1
        if (pointer > 1000000) {
            print(pointer)
            break
        }
    }
    print(pointer)
}
```

For (cos(x) / 30) + (44 * x / 45) all iterations converge for 0, 0.5, 0.7. We notice that the starting point with 0.7 has the least iterations, this
is to b expected since the root is at 0.914...

## Chapter 4.1.2 question 3

```{r}
twin_primes <- list()
for (i in (0:1000))
{
    for (x in (i:1000)) {
        if ((x == i + 2)) {
            switch <- TRUE
            for (z in (2:(x - 1))) {
                if (x %% z == 0) {
                    switch <- FALSE
                    break
                }
            }
            for (z in (2:(i - 1))) {
                if (i %% z == 0) {
                    switch <- FALSE
                    break
                }
            }
            if (switch == FALSE) {
                break
            }
            twin_primes[[length(twin_primes) + 1]] <- c(x, i)
        }
    }
}
print("All the values")
twin_primes

print("The length of the list, counting the numbers of pairs")
length(twin_primes)

print("The length of the list, counting the numbers of individual elements")
length(twin_primes) * 2

print("the 10th pair")
twin_primes[[10]]

print("the 15th pair")
twin_primes[[15]]
```

We create a for loop that compares number i ranging from 1 to 1000, which checks if the equation x = i + 2 is satisfied for x which takes all the subsequent numbers after i to avoid permutations of the same 
elements. We then check if x and y are prime number by dividing them by all numbers from 2 to n-1. If they are not a prime te 'switch' variable will be set to false and the second nested loop will break.

## Chapter 4.1.2 question 4
```{r}
GIC <- function(P, n) {
    if (P <= 0 || n <= 0 || n %% 1 != 0) {
        stop("error, principal cannot be negative, n cannot be negative, n must be integer")
    }
    if (n <= 3) {
        i <- 0.04
    } else {
        i <- 0.05
    }
    return(P * ((1 + i)^n - 1))
}
GIC(1000, 2)
GIC(1000, 20)
```

To determine wether we use an interest of 0.04 or 0.05 we set i to 0.04 if the condition of n smaller or equal to 3 is respected. If not, i is equal to 0.05. Once the i is decided we compute
the formula with the given i over n time periods