---
title: "Assignment 3"
author: 'Terrance Liang 40158057'
output: 'pdf_document'

---

## Chapter 2.6 problem 1 

```{r}
# creating the data we want to perform operations on
solar.radiation <- c(11.1, 10.6, 6.3, 8.8, 10.7, 11.2, 8.9, 12.2)
# creating a function to avoid repetitino
descriptive_stats <- function(input, string_v) {
    print(string_v)
    print(mean(input))
    print(median(input))
    print(range(input))
    print(var(input))
}
# displaying the desacriptive statistics of solar.radiation
descriptive_stats(solar.radiation, "solar radiation--")
sr10 <- solar.radiation + 10
# displaying the descriptive statistics of sr10
descriptive_stats(sr10, "sr10-------------")
srm2 <- solar.radiation * -2
# displaying the descriptve statistics of srm2
descriptive_stats(srm2, "srm2-------------")
# histogram plotting of all the the variables
hist(solar.radiation)
hist(sr10)
hist(srm2)
```
for c) we notice that there is no change in the variance, but that there is change in the 
other descriptive measures. The increase of 10 in the range, the mean and median is explained
 by the fact that all measures are increased by 10, thus it follows that all prior descriptive measures are also increased by the same amount.
 The variance does not change because the measure of dispersion from the mean is still the same, this can 
 be seen by the fact that the histogram is the same, but that it is simply shifted by 10 units.

 for d) we notice that all values change, the mean is now -2 times the prior mean, the median has also changed because our range of 
 values have changed, thus the new median is calculated from a range that is now negative and wider. The range is
 wider and negative, and the variance is now wider because our values are more spread out. The -2 operator
 spreads the values.


## Chapter 2.6 problem 3

```{r}
# we first create 2 vectors of the required vector, pairwise maximum is then computed 
# between those 2 vectors then we sum all the pairwise max

sum(pmax(2^(1:15),(1:15)^3))
```

## Chapter 2.8 problem 2

```{r}
data <- USArrests
data
# number of rows
nrow(data)
# number of columns
ncol(data)
# median of every column
lapply(data, median)
# murder per capita average 77%, we first select the rows which have more than 77 murder per 
# capita, then only select the murder column, then we apply the mean for those values
lapply(data[data$UrbanPop > 77, ]["Murder"], mean)
# murder per capita average 50%
lapply(data[data$UrbanPop < 50, ]["Murder"], mean)
# new smaple of data, we select a sampe of the row indexes without replacement of size 12 and use
# it as conditional indexing for data
newData <- data[sample((1:nrow(data)), size = 12, replace = FALSE), ]
newData
```

## Chapter 2 problem 3

```{r}
data <- chickwts
# boolean indexing of chickweights with weight baove 300
chickwts300 <- chickwts[chickwts$weight > 300, ]
# boolean idnexing of chicweights who were fed with linseed
chickwtsLinseed <- chickwts[chickwts$feed == "linseed", ]
# everage weight of chickes fed linseed
mean(chickwtsLinseed$weight)
# average weight of chicks not fed linseed
mean(chickwts[chickwts$feed != "linseed", ]$weight)
```

## Chapter 2 problem 6

```{r}
# curve of the function from 0 to 3, changed the x and y range that is observable from the plot
curve(expr = 3 * x + 2, from = 0, to = 3, xlim = c(0, 6), ylim = c(-30, 30), ylab = "y")
# graph from the function from 3 to 6, added argument add= true to ad donto an existing plot
curve(expr = 2 * x - (0.5 * x^2), from = 3, to = 6, add = TRUE)
```

## Chapter 2 problem 7

```{r}
# sampling a die roll 1000000 times
dieRolls <- sample((1:6), size = 1000000, replace = TRUE)
# creating a factor of dierolls
dieRollsFactor <- factor(dieRolls)
# assigning new levels to the dierollsfactors
levels(dieRollsFactor) <- c(
    "One", "Two", "Three", "Four", "Five",
    "Six"
)
# casting dierollsfactor as characters
dieRollsChar <- as.character(dieRollsFactor)

# creating tables of the three varibales 
table(dieRolls)
table(dieRollsFactor)
table(dieRollsChar)

# checking the time sit takes to convert each varibale into a table
system.time(table(dieRolls))
system.time(table(dieRollsFactor))
system.time(table(dieRollsChar))

# creating a file with the specific varibale saved inside them 
dump("dieRolls", "dieRolls.R")
dump("dieRollsFactor", "dieRollsFactor.R")
dump("dieRollsChar", "dieRollsChar.R")
# checking file size
file.info("dieRolls.R")
file.info("dieRollsFactor.R")
file.info("dieRollsChar.R")
```
for d) each data set has the same amount of occurences for its repsective number. 1 has the same number of occurences
as 'one'. The main difference is the order of the columns, dieRolls is ordered numerically 1 through 6
in ascending order. dierollsfactor is ordered numerically and ascending by the string's numerical 
interpretation of the number. dieRollsChar is ordered by the character's value, thus F comes before 
O, thus Five comes before One. For f) in ascending order the slowest to fastest table were dieRollsFactor,
dieRollsChar,dieRolls. The third column shows the elapsed time of the process. For f) in ascending order
 dieRolls, dieRollsFactor, dieRollsChar.