---
title: "Assignment 1"
author: 'Terrance Liang 40158057'
output: 'pdf_document'
---

## problem 1 a

Here we assign to a variable a function performing the calculations asked based on an input and which returns the output after the calculations
```{r}
fun <- function(input) {
    return(cos(input * exp(1))^3 + log(5 * input, 3) - atan(input))
}

fun(1)
```

## problem 1 b

we create a vector with 10^6+1 points all evenly spaced ranging from 1 to 2, these correspond to the parameters et in the "seq" function
```{r}
grid <- c(seq(from = 1, to = 2, length.out = 10^6 + 1))
```

## problem 1 c

we use the sample function on the grid variable we created, this samples n values from the vector without replacement and then assigns it
to an attribute of of subgrid variable
```{r}
m <- c(10, 100, 1000)

subgrid.1 <- sample(grid, size = m[1], replace = FALSE)
subgrid.2 <- sample(grid, size = m[2], replace = FALSE)
subgrid.3 <- sample(grid, size = m[3], replace = FALSE)
```

## problem 1 d

we compute the mean of the output of the functions using each subgrid's attribute as input and store it inside another variable's attribute in 
the form of a vector of length 3.
```{r}
eval.1 <- fun(subgrid.1)
eval.2 <- fun(subgrid.2)
eval.3 <- fun(subgrid.3)

monte.carlo <- c(mean(eval.1), mean(eval.2), mean(eval.3))

print(mean(monte.carlo))
```

## problem 1 e

we compute the integral of the function
```{r}
integral <- 0.479199

integral_full <- integrate(fun, 1, 2)

print(integral_full)

print(monte.carlo - integral)
```

We notice that the integral of the function is in the neighborhood of the samples taken. The differences between the mean value of every column is 
very minimal with the integral as well as the difference between the entire mean of the vector with the integral. The bigger the sample the smaller the 
difference.

## chapter 2.1 problem 3

we compute the total interest earned at every period and the ones prior to the current period
```{r}
investment <- 2000 * ((1 + 0.03)^(1:30)) - 2000
print(investment)
```

## chapter 2.1 problem 6

we compute the area of pi for various radius and list them
```{r}
radius <- pi * (3:100)^2
print(radius)
```

## chapter 2.3 problem 1

we compute the geometric sum until n the long and the fast way and compare the differences
```{r}
geometric_sum_long <- function(n, r) {
    sum <- 0
    for (x in (0:n))
    {
        sum <- sum + r^x
    }

    return(sum)
}

geometric_sum_short <- function(n, r) {
    return((1 - r^(n + 1)) / (1 - r))
}

for (x in seq(10, 40, 10))
{
    print(geometric_sum_long(x, 1.08))
    print(geometric_sum_short(x, 1.08))
}

for (x in seq(10, 40, 10))
{
    print(geometric_sum_long(x, 1.06))
    print(geometric_sum_short(x, 1.06))
}
```

## chapter 2.3 problem 2
we perform the geometric sum for another ratio but for the same range
```{r}
sum_vector <- c(geometric_sum_short((1:100), 1.08))
print(sum_vector)
```

### chapter 2.3 problem 8
for vector 1 we create a sequence ranging from 0 to 4 for which every value of
that sequence is individually and subsequently repeated based at it's corresponding index indicated by rep(5,5)

for vector 2 we generate a sequence from 1 to 5 which is repeated 5 times subsequently but not individually
```{r}
vector_1 <- rep(seq(0, 4), rep(5, 5))
print(vector_1)

vector_2 <- rep(seq(1, 5), 5)
print(vector_2)
```

## chapter 2.3 problem 9

we generate a sequence from 1 to 5 which is subsequently repeated and increment every interval of of 5 values by 0 then the next 5 values by 1 and so on
```{r}
vector_3 <- rep(seq(1, 5), 5) + rep(seq(0, 4), rep(5, 5))
print(vector_3)
```