---
title: "Assignment 4"
author: 'Terrance Liang 40158057'
output: 'pdf_document'

---

## Chapter 3.1 question 1

a)
```{r}
par(mfrow=c(2,2))
hist(islands, breaks = "Sturges", main = "Sturges")
hist(log(islands), breaks = "Sturges", main = "Sturges on log scale")
hist(islands, breaks = "Scott", main = "Scott")
hist(log(islands), breaks = "Scott", main = "Scott on log scale")
```

b)
```{r}
par(mfrow=c(1,2))
qqplot( rnorm(48),log(1000 * islands, 10))
qqplot( rnorm(48),islands)
```

We construct a QQplot using the untouched island data and the log scaled island data. In the first case,
it has the most similarities with the 4th picture titled "D is skewed to the right". For the log scaled 
data, it also has the most similarities with the 4th picture. In both cases, because of the heavy rightward 
skewness, we expect the orignal data to follow a distribution with some similarities to the exponential one.

c)
```{r}
par(mfrow=c(1,2))
boxplot(x=islands, main = "Original scale")
boxplot(x=log(islands), main = "Log scale")
```

d)

```{r}
par(mfrow=c(1,2))
dotchart(x=islands,main="Original scale",cex = 0.3, cex.main = 2)
dotchart(x=log(islands),main="Log scale",cex = 0.3, cex.main = 2)
```
Using the log scale we notice that differences are more easily visualisable as compared to 
the orginal scale where most values are all at the inital line. Thus the log scale enables better 
visualisation of differences as opposed to the orignal one.

e)
Different graphics have different uses, thus there is not a better overall graphic for displaying data,
QQplot give insight on data normality/skewness but a dotchart gives a mean for comparison. Using one  
for the purpose of the other would provide no isight on the matter we are looking for. Thus the better
graphic depends on the use and the purpose we are looking for.

In terms of what is the better graphic amongst the ones we have used for its specific purpose, 
the sturges histogram on log scale for comparing frequency, the log scaled dot chart for investigasting skewness
and normality, then the log scaled box plot for identifyig quartiles and outliers, and finally
the log scaled dot chart for comparing differences.

## Chapter 3.1 question 3

a)
```{r}
plot(pressure ~ temperature, data = pressure)
```
The variables seem to be non linearly related because of the evident curve in the data points,
b)
```{r}
residuals <- with(pressure, pressure - (0.168 + 0.007 * temperature)^(20 / 3))
qqplot(rnorm(length(residuals)), residuals)
```
Because the plot does not represent a straight line when plotted against a normal distribution, 
the residuals are therefore not normally distributed. We also notice that the plot increases at a faster 
rate a the values increase on the horizontal axis, therefore there is a skewed to the right distribution.

c)
```{r}
plot(pressure^(3/20) ~ temperature, data = pressure)
```
After transformation, there is now a clear linear relationship between the temperature and the Transformed
pressure values.
d)
```{r}
residuals <- with(pressure, pressure^(3/20) - (0.168 + 0.007*temperature)^(20/3))
Normal <- rnorm(length(residuals))
qqplot(Normal, residuals)
```
The QQplot shows that the residuals do not follow a normal distribution due to the non linearity of the 
plot. The rate of increase decreases as the values on the horizontal axis grow. Thus it follows a left skewed distribution.

## Chapter 3.3 question 3

```{r}
par(list = list(mfrow = c(3, 2), mar = c(2.5, 2.1, 0.1, 0.1)))
Z <- (rnorm(1000))^2
hist(Z, main = "")
hist(Z, breaks = "Freedman-Diaconis", main = "")
plot(density(Z), main = "")
boxplot(Z)
qqnorm(Z, main = "")
qqline(Z)
ts.plot(Z)
```
Comparing our modified plots with the plots in figure 3.18 we notice that the histogram is now
skewed to the right. The histogram using the freedman-diaconis breaks has also now become right 
skewed. The density plot have also now become right skewed. the box plot had its lower tail shortened, more outliers
and it's overall box size reduced. the QQ plot is now not normal anymore but right skewed instad, the tsplot now has 
higher peaks. It is also important to note that all the axis which had negatiev values are now 
only located in the positive parts of their respectiev axis due to the square operator.

## Chapter 3.3 question 4
```{r}
par(list = list(mfrow = c(3, 2), mar = c(2.5, 2.1, 0.1, 0.1)))
Z <- EuStockMarkets
hist(Z, main = "")
hist(Z, breaks = "Freedman-Diaconis", main = "")
plot(density(Z), main = "")
boxplot(Z)
qqnorm(Z, main = "")
qqline(Z)
ts.plot(Z)
summary(Z)
```
We notice that it is right skewed, there are presence of outliers past the upper whisker and the upper whisker is very large as opposed to the 
lower one. The datasets are not normal. The stock with the all time high is SMI, the lowest low is DAX, the highest mean is FTSE.

The histograms and the density plots do not show further insight, we are able to discern that it is right sweked based on the
QQplot. The first 3 plots only show frequency at certain points and no additional insight is given. The last 3 plots are useful
since the box plot shows outlier values for every stock perhaps showing overvaluation and undervaluation values. The
QQ plot shows that the values are right skewed showig that the data is not normal, the time series plot
graphs the values of each stock with respect to time, which gives a mean of comprison across the different
stocks at the chosen point in time.

## Chapter 3 question 2

a)
```{r}
plot(pressure ~ temperature, pressure)
curve((0.168 + 0.007 * x)^(20 / 3), from = 0, to = 400, add = TRUE)
```

b)
```{r}
plot(pressure^(3 / 20) ~ temperature, pressure)
abline(lm(pressure^(3 / 20) ~ temperature, pressure))
```
The plot is now shows that both variables are linearly related
c)
```{r}
plot(pressure^(3 / 20) ~ temperature, pressure)
abline(lm(pressure^(3 / 20) ~ temperature, pressure))
title("Transformed pressure \n value vs temperature")
```

d)
```{r}
par(mfrow = c(2, 1))
plot(pressure ~ temperature, pressure)
curve((0.168 + 0.007 * x)^(20 / 3), from = 0, to = 400, add = TRUE)
title("pressure \nvalue vs temperature")
plot(pressure^(3 / 20) ~ temperature, pressure)
abline(lm(pressure^(3 / 20) ~ temperature, pressure))
title("Transformed pressure \nvalue vs temperature")

par(mfrow = c(1, 2))
plot(pressure ~ temperature, pressure)
curve((0.168 + 0.007 * x)^(20 / 3), from = 0, to = 400, add = TRUE)
title("pressure \nvalue vs temperature")
plot(pressure^(3 / 20) ~ temperature, pressure)
abline(lm(pressure^(3 / 20) ~ temperature, pressure))
title("Transformed pressure \nvalue vs temperature")
```