---
title: "Assignment 6"
author: 'Terrance Liang 40158057'
output: 'pdf_document'
---

## Chapter 4.1.3 question 1

```{r}
# assignment of initial fibonnaci values 
Fib1 <- 1
Fib2 <- 1

# create a vector of the current fibonnaci values
Fibonacci <- c(Fib1, Fib2)

# while loop to stop at fibbonacci value exceeding 300
while (Fib2 < 300) {
#concatenates all past fib values with the most recent one  
    Fibonacci <- c(Fibonacci, Fib2)
# computes the newest fib value 
    Fib2 <- Fib1 + Fib2
# assigns the last value to fib 1 to compute fib 2 after  
    Fib1 <- Fibonacci[length(Fibonacci)]
}
print(Fibonacci)
```

## Chapter 4.1.3 question 2


```{r}
# creation of initial fibonnaci vales
Fibonacci <- c(1, 1, 1)
# while loop to validate until the last 2 values if their sum is less than 300
while ((Fibonacci[length(Fibonacci)] + Fibonacci[length(Fibonacci) - 1]) < 300) {
    # appends the last 2 values to the fibonnaci sequence
    Fibonacci <- append(Fibonacci, (Fibonacci[length(Fibonacci)] + Fibonacci[length(Fibonacci) - 1]))
}
print(Fibonacci)

```


## Chapter 4.1.3 question 3

```{r}
Fibonacci <- c(1, 1, 1)
while ((Fibonacci[length(Fibonacci)] + Fibonacci[length(Fibonacci) - 1]) < 1000000) {
    Fibonacci <- append(Fibonacci, (Fibonacci[length(Fibonacci)] + Fibonacci[length(Fibonacci) - 1]))
}
length(Fibonacci)
```

We repeat the code from the last problem and we output the length og the resulting vector 

## Chapter 4.1.3 question 4
```{r}
# for the sake of comparison we have a 0 as our initial vector element to be able to compare 
# inside the while loop
po <- c(0, 0.006)

# check the absolute difference between the last 2 elements of the vector
while (abs(po[length(po)] - po[length(po) - 1]) > 0.000001) {
    # add a new element to the vector which follows the provided interest formula 
    po <- c(po, (1 - (1 + po[length(po)])^(-20)) / 19)
}
print(po)
```

```{r}
for (i in c(10, 1, 0.1, 0.01, 0.001, 0.0001)) {
    po <- c(0, i)

    while (abs(po[length(po)] - po[length(po) - 1]) > 0.000001) {
        po <- c(po, (1 - (1 + po[length(po)])^(-20)) / 19)
    }
    print(po[length(po) - 1])
    print(po[length(po)])
    print(" ")
}
```
If we use other starting guesses there will simply be more values outputted until we reach the absolute
 difference of 0.000001, however, for bigger numbers we notice there is a trend for which a constant number of outputs are produced
 
## Chapter 4.1.3 question 5

```{r}
po <- c(0, 0.006)
count <- 0
while (abs(po[length(po)] - po[length(po) - 1]) > 0.000001) {
    count <- count + 1
    po <- c(po, (1 - (1 + po[length(po)])^(-20)) / 19)
}
print(count)
```
We simply add a counter in the while loop that increases by 1 for every iteration

```{r}
for (i in c(10, 1, 0.1, 0.01, 0.001, 0.0001)) {
    po <- c(0, i)
    count <- 0
    while (abs(po[length(po)] - po[length(po) - 1]) > 0.000001) {
        count <- count + 1
        po <- c(po, (1 - (1 + po[length(po)])^(-20)) / 19)
    }
    print(count)
}
```

we notice that the number of iterations is constant for bigger starting points

## Chapter 4.1.4 question 4

```{r}
D(expression((x - 3)^4 + 7 * (x - 2)^2 + x), "x")
newton <- c(0, 3)
while (abs(newton[length(newton)] - newton[length(newton) - 1]) > 0.01) {
    # assigning the last output to x for simplicity
    x <- newton[length(newton)]
    # adding the new output to the sequence of newton outputs
    newton <- c(newton, x - (((x - 3)^4 + 7 * (x - 2)^2 + x - 2.82) / (4 * (x - 3)^3 + 7 * (2 * (x - 2)) + 1)))
}
print(newton[length(newton)])
```
We derived the equation and solving for f'(x) = 0 gives us 2.12 has a slope of 0 at that point. We then find that at 2.12 f(x)=2.82 thus 
for the sake of using newton's method we shift down by 2.82 so that there is a 0 to the function.

## Chapter 4.1.4 question 5

The function f (x) = (5x - 3)/(x - 1) has a single zero, since we are able to get (5x - 3 -2 +2)/(x - 1) = 5 + 2/(x-1) which is a rational function.
Equalling it to zero gives that the zero is -2/5 +1.

```{r}
D(expression(5 + 2 / (x - 1)), "x")
for (i in c(0.5, 0.75, 0.2)) {
    newton <- c(0, i)
    while (abs(newton[length(newton)] - newton[length(newton) - 1]) > 0.00000001) {
        # assigning the last output to x for simplicity
        x <- newton[length(newton)]
        # adding the new output to the sequence of newton outputs
        newton <- c(newton, x - ((5 + 2 / (x - 1)) / (-(2 / (x - 1)^2))))
    }
    print(newton)
}
```
We notice that the first three converges to 0.6, 0.6 and 1 respectively, the last initial point gives and error on further inspection in the following
block of code we notice that it in fact diverges to inficity
```{r}
# newton <- c(0, 1.25)
# while (abs(newton[length(newton)] - newton[length(newton) - 1]) > 0.00000001) {
#     # assigning the last output to x for simplicity
#     x <- newton[length(newton)]
#     # adding the new output to the sequence of newton outputs
#     newton <- c(newton, x - ((5 + 2 / (x - 1)) / (-(2 / (x - 1)^2))))
#     print(newton[length(newton)])
# }
```

## Chapter 4.1.4 question 6

Seperating the equation into two parts yields (x^2 - 6 * x + 9) and e^(-x), since the other side equals 0, this is allowed since dividing by either or will
leave the other side intact. the first part has a single 0 at x= 3, the second one has no zeros since ln(0)=x is unsolvable 
```{r}
D(expression((x^2 - 6 * x + 9) * exp(-x)), "x")
for (i in c(3.2, 2.99, 3.01)) {
    newton <- c(0, i)
    while (abs(newton[length(newton)] - newton[length(newton) - 1]) > 0.00000001) {
        # assigning the last output to x for simplicity
        x <- newton[length(newton)]
        # adding the new output to the sequence of newton outputs
        newton <- c(newton, x - (((x^2 - 6 * x + 9) * exp(-x)) / ((2 * x - 6) * exp(-x) - (x^2 - 6 * x + 9) * exp(-x))))
    }
    print(newton)
}

```
We notice that the last three converges to 3 respectively, the first initial point gives and error on further inspection in the following
block of code we notice that it in fact gives a nan value which we cannot continue out iteration from on the first value.
```{r}

# newton <- c(0, 3)
# while (abs(newton[length(newton)] - newton[length(newton) - 1]) > 0.00000001) {
#     # assigning the last output to x for simplicity
#     x <- newton[length(newton)]
#     # adding the new output to the sequence of newton outputs
#     newton <- c(newton, x - (((x^2 - 6 * x + 9) * exp(-x)) / ((2 * x - 6) * exp(-x) - (x^2 - 6 * x + 9) * exp(-x))))
#     print(newton)
# }

```

## Chapter 4.1.5 question 2.cannot
```{r}
Eratosthenes <- function(n) {
    # Print prime numbers up to n (based on the sieve of Eratosthenes)
    if (n >= 2) {
        sieve <- seq(2, n)
        primes <- c()
        while (length(sieve) > 0) {
            p <- sieve[1]
            primes <- c(primes, p)
            sieve <- sieve[(sieve %% p) != 0]
            if (p >= sqrt(n)) {
                break
            }
        }
        primes <- c(primes, sieve)
        return(primes)
    } else {
        stop("Input value of n should be at least 2.")
    }
}

Eratosthenes(10)
```

We create as break condition p >= sqrt(n) which says that if this is true all remaining numbers in sieve are prime. Thus for n =10 the sieve number 10
need not be checked for prime, it is automatically prime given the condition.